---
title: "Checking discrepancies"
author: "Johanna Jantzen"
date: "November 18, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spectrolab)
library(tidyr)
library(dplyr)
library(lubridate)
library(reshape2)

```

Etienne download versus my fulcrum download differences
```{r fulcrum downloads}

#read in spectral data object

Etienne_data <- read.csv("https://web.fulcrumapp.com/shares/5c5290b884ab3734.csv") # Fulcrum data share link

Shan_spectra <- readRDS("./data/all_spectra.rds")

#get metadata object from spectra
metadata_Shan <- meta(Shan_spectra)

#get species list from spectral data
species_Shan <- meta(Shan_spectra, "species", simplify = TRUE)

#get metadata from additional metadata file (downloaded separately from fulcrum)
Johanna_data <- read.csv("./data/metadata/leaf_spectra.csv", stringsAsFactors = FALSE)

```

Checking for differences in my and Etienne's metadata

``` {r compare}

#check which sample IDS are not in my sample IDs
Johanna_missing <- Etienne_data[-which(Etienne_data$sample_id %in% Johanna_data$sample_id),]

Etienne_missing <- Johanna_data[-which(Johanna_data$sample_id %in% Etienne_data$sample_id),]



nrow(Johanna_missing)
nrow(Etienne_missing)

```
Mine is missing 461 and Etienne's is missing 0 samples 

``` {R which mising}


#there are no NA scientific names in Etienne's data
Etienne_data[which(is.na(Etienne_data$scientific_name)),]

#there are no NA scientific name in Johanna's data
Johanna_data[which(is.na(Johanna_data$scientific_name)),]

#there are 126 names in Etienne's data
unique(Etienne_data$scientific_name)

#there are 91 names in Johanna's data therefore 35 species names are missing in Johanna's metadata
unique(Johanna_data$scientific_name)

#there are samples of 71 species missing from Johanna's metadata that are in Etienne's (maybe not unique names)
unique(Johanna_missing$scientific_name)

#these missing samples include samples from four categories of data
unique(Johanna_missing$status)

#there are samples from 8 projects missing in my metadata that are in Etienne's
unique(Johanna_missing$project)

#which project are the deleted ones missing from? 3: CABO-test, 2018-Hacker-PhD-UBC and 2019-Blanchard-MSc-UdeM
unique(Johanna_missing$project[which(Johanna_missing$status == "deleted")])

#which project are the NOT deleted ones missing from? 8 projects
unique(Johanna_missing$project[-which(Johanna_missing$status == "deleted")])

#Verified samples are missing from the CABO-test project
unique(Johanna_missing$project[which(Johanna_missing$status == "verified")])

#Pending samples are missing from CABO-test and 2019-Pardo-MSc-UdeM
unique(Johanna_missing$project[which(Johanna_missing$status == "pending")])

#Submitted samples are missing from 7 projects
unique(Johanna_missing$project[which(Johanna_missing$status == "submitted")])

#Johanna's data is missing all 2019-Pardo while Etienne's has 71 samples from 2019-Pardo
"2019-Pardo-MSc-UdeM" %in% unique(Johanna_data$project)
"2019-Pardo-MSc-UdeM" %in% unique(Etienne_data$project)

length(Etienne_data$sample_id[which(Etienne_data$project == "2019-Pardo-MSc-UdeM")])

#Are all the missing samples from Pardo? No, but samples of 71 species are (71 of the 461 are)
length(Johanna_missing$sample_id[which(Johanna_missing$project == "2019-Pardo-MSc-UdeM")])
length(Johanna_missing$sample_id[-which(Johanna_missing$project == "2019-Pardo-MSc-UdeM")])
nrow(Johanna_missing)

```

Now to compare Etienne's data with the spectral data

```{r spectral comparison}

# Etienne_data 
# 
# Shan_spectra 
# 
# metadata_Shan 
# 
# species_Shan 
# 
# Johanna_data

#which samples are missing from full metadata 
Etienne_missing_from_Shan <- metadata_Shan[-which(metadata_Shan$sample_id %in% Etienne_data$sample_id),]
Shan_missing_from_Etienne <- Etienne_data[-which(Etienne_data$sample_id %in% metadata_Shan$sample_id),]
Johanna_missing_from_Shan <- metadata_Shan[-which(metadata_Shan$sample_id %in% Johanna_data$sample_id),]

#200 samples missing from Etienne metadata in Shan spectral data
nrow(Etienne_missing_from_Shan)

#626 samples missing from Johanna metadata in Shan spectral data
nrow(Johanna_missing_from_Shan)

#which projects are missing - 2017-Dessain-MSc missing from Etienne data
unique(Etienne_missing_from_Shan$project)
unique(Shan_missing_from_Etienne$project)
unique(Johanna_missing_from_Shan$project) # missing from 8 projects

#have species IDs for samples missing metadata but not other metadata
is.na(Etienne_missing_from_Shan$species)

Johanna_missing_from_Shan$project[which(is.na(Johanna_missing_from_Shan$species))] #included somes samples which had no species ID, from Pardo and SWA-Warren projects

```

It seems that the main issues (now that I have Etienne's data with more of the discrepancies resolved), is that the metadata is missing for the 2017-Dessain-MSc project. 
For the data availble for me (Johanna) to download from the Fulcrum website, there was more information missing compared to the spectral data, including for samples from 8 projects (with all of the 2019-Pardo project missing), and including samples without species IDs in Shan's spectral data. 
