---
title: "Summarizing data"
author: "Johanna Jantzen"
date: "November 18, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spectrolab)
library(tidyr)
library(dplyr)
library(lubridate)
library(reshape2)
library(ggplot2)
library(forcats)
library(taxize)
library(jsonlite)
library(purrr)
library(tibble)
library(stringr)

```

``` {r create functions,  include=FALSE}

#make presence_absence function
presence_absence <-function(data){
  tnt <- as.data.frame(data)
  # Create an ID column to hold the row index
  tnt$ID <- seq.int(nrow(tnt))
  out=dcast(tnt, ID ~data, length)
  
  # Get rid of the ID column from the matrix
  drops=c("ID")
  out=out[ , !(names(out) %in% drops)]
  
  return(out)
}

#make splitting species names function
splitting_species_names <- function(species_names_dataframe){
  sample_meta_names <- species_names_dataframe %>% 
    separate(PL.scientific_name, c("genus", "specific_epithet", "extra0", "extra1", "extra2", "extra3", "extra4", "extra5", "extra6", "extra7", "extra8", "extra9", "extra10"), sep = " ", fill = "right", remove = FALSE)
  
  #deal with subspecific names
  sample_meta_names$subspecific <- NA
  sample_meta_names$subspecific[which(sample_meta_names$extra5 == "australis")] <- paste0(sample_meta_names$extra4[which(sample_meta_names$extra5 == "australis")], "_", sample_meta_names$extra5[which(sample_meta_names$extra5 == "australis")])
  
  sample_meta_names$subspecific[which(sample_meta_names$extra2 == "angustifolia")] <- paste0(sample_meta_names$extra1[which(sample_meta_names$extra2 == "angustifolia")], "_", sample_meta_names$extra2[which(sample_meta_names$extra2 == "angustifolia")])
  sample_meta_names$subspecific[which(sample_meta_names$extra2 == "vaginatum")] <- paste0(sample_meta_names$extra1[which(sample_meta_names$extra2 == "vaginatum")], "_", sample_meta_names$extra2[which(sample_meta_names$extra2 == "vaginatum")]) 
  
  sample_meta_names$subspecific[which(sample_meta_names$extra1 == "spissum")] <- paste0(sample_meta_names$extra0[which(sample_meta_names$extra1 == "spissum")], "_", sample_meta_names$extra1[which(sample_meta_names$extra1 == "spissum")])
  sample_meta_names$subspecific[which(sample_meta_names$extra1 == "strigosus")] <- paste0(sample_meta_names$extra0[which(sample_meta_names$extra1 == "strigosus")], "_", sample_meta_names$extra1[which(sample_meta_names$extra1 == "strigosus")]) 
  sample_meta_names$subspecific[which(sample_meta_names$extra1 == "rugosa")] <- paste0(sample_meta_names$extra0[which(sample_meta_names$extra1 == "rugosa")], "_", sample_meta_names$extra1[which(sample_meta_names$extra1 == "rugosa")])
  return(sample_meta_names)
}


```
##Summary of Data

In this document, I summmarize the data available from the CABO project. 

I am using Fulcrum metadata downloaded on November 17, 2020 using the following link. 


```{r download metadata}

#read in metadata object
#metadata_fulcrum <- read.csv("https://web.fulcrumapp.com/shares/5c5290b884ab3734.csv", stringsAsFactors = FALSE) # Fulcrum data share link (first edit column names for reading)
metadata_fulcrum <- read.csv("./data/exported_raw/leaf_spectra.csv", stringsAsFactors = FALSE)

```


The spectral data available was compiled by Shan Kothari (received October 21, 2020). 

```{r read in spectral data}

#this is the preliminary version (cleaned by Shan)
spectral_data <- readRDS("./data/all_spectra.rds")

#get the metadata from the spectra object
metadata_Shan <- meta(spectral_data)

```

I will first reduce the metadata to those samples with spectra in Shan's spectral data file. Those samples missing from the spectral file are saved in a list. 

``` {r match datasets}

#keep only the metadata from fulcrum for samples in Shan's spectral file
matched_meta <- metadata_fulcrum[which(metadata_fulcrum$sample_id %in% metadata_Shan$sample_id),]
nrow(matched_meta)

#those spectra without metadata
missing_spectra <- metadata_fulcrum[-which(metadata_fulcrum$sample_id %in% metadata_Shan$sample_id),]
nrow(missing_spectra)

#there are some submitted samples without spectra (only 19?)
unique(missing_spectra$status)
#missing_spectra[which(missing_spectra$status == "submitted"),]

#those samples without spectra
missing_metadata <- metadata_Shan[-which(metadata_Shan$sample_id %in% metadata_fulcrum$sample_id),]
nrow(missing_metadata)

#Exclude spectra from Dessain because no metadata and because they used a different instrument to measure spectra
unique(missing_metadata$project)

#save missing data objects
saveRDS(missing_spectra, "./data/tidy/fulcrum_missing_spectra.rds")
saveRDS(missing_metadata, "./data/tidy/spectra_missing_fulcrum.rds")

```


With the matched dataset, I merge the leaf spectra (only those whose status is submitted), bulk sample, and plant dataframes. I check for duplicate sample IDs and errors in matching plant ID to sample ID. 

``` {r match species names}

#make sure no missing IDs for metadata
still_missing_species_IDS <- matched_meta$sample_id[which(is.na(matched_meta$species))]

#make sure only submitted samples - remove statuses
#nrow(matched_meta)
status_matched_meta <- matched_meta[which(matched_meta$status == "submitted"),]
#nrow(status_matched_meta)

#need to match plant ID with sample ID because taxonomy of leaf spectra metadata is not updated (as a rule)

#read plants data
plants <- read.csv("./data/exported_raw/plants.csv", stringsAsFactors = FALSE)

#read bulk sample data to merge plants and leaf spectra data
bulk_samples <- read.csv("./data/exported_raw/bulk_leaf_samples.csv", stringsAsFactors = FALSE)

#retrieve plant ids for samples in leaf spectra (separately)
plant_id_for_spectra <- bulk_samples[which(bulk_samples$sample_id %in% status_matched_meta$sample_id),]
nrow(plant_id_for_spectra) #2051
length(status_matched_meta$sample_id) #2052 - one more sample than have plant ids
#there is one sample that has a unqiue sample_id but no plant_id

#checking for duplicate sample_ids
unique(status_matched_meta$sample_id %in% bulk_samples$sample_id)
status_matched_meta[which(duplicated(status_matched_meta$sample_id) == TRUE),c(9,10,12,13,19,20,21,22)] #sample_id = 44682364
#status_matched_meta[which("44682364"%in% status_matched_meta$sample_id),]
bulk_samples[which(bulk_samples$sample_id == "44682364"),c(9,10,12,13,21,22,23,26,30)] #plant_id = 44226204; sample_id = 44682364
status_matched_meta[which(status_matched_meta$sample_id == "44682364"),c(4,9,10,12,13,19,20,21,22,23,26,30)] #two samples with different gps but same sample_id
#bulk_samples[which(bulk_samples$sample_id == "44682364"),] #two samples with different gps but same sample_id, and one plant id (for the sample id in bulk samples)
nrow(status_matched_meta) #2052

#remove sample without plant id
status_matched_meta <- status_matched_meta[-which(status_matched_meta$sample_id == "44682364"),]
nrow(status_matched_meta) #2050

#there is no missing plant id info in bulk samples for spectra after removing the duplicate sample ID
missing_plant_id_for_spectra <- status_matched_meta$sample_id[-which(status_matched_meta$sample_id %in% bulk_samples$sample_id)]

#get species IDs for plant ids in spectra
species_ids_for_spectra <- plants[which(plants$plant_id %in% plant_id_for_spectra$plant_id),]
nrow(species_ids_for_spectra)

# #quite a few samples have duplicate plant_ids (as expected)
# plant_id_for_spectra[which(duplicated(plant_id_for_spectra$plant_id) == TRUE),]

#make new dataframe for matching species ids and plant_ids

#first change name of columns to match origin
colnames(status_matched_meta) <- paste0("LS.", names(status_matched_meta))
colnames(bulk_samples) <- paste0("BS.", names(bulk_samples))
colnames(plants) <- paste0("PL.", names(plants))
#colnames(plants)

#first join bulk samples by sample id
species_reconc <- merge(status_matched_meta, bulk_samples, by.x = "LS.sample_id", by.y = "BS.sample_id", all.x = TRUE)
nrow(species_reconc)

length(unique(species_reconc$BS.plant_id)) #1804
length(unique(species_reconc$LS.sample_id)) #2050
#species_reconc$LS.sample_id

#then add plant ids
species_reconc <- merge(species_reconc, plants, by.x = "BS.plant_id", by.y = "PL.plant_id", all.x = TRUE) #those colnames without a specifier are from the plants dataframe
nrow(species_reconc)
#why does this step add a sample
length(unique(species_reconc$BS.plant_id)) #1804
length(unique(species_reconc$LS.sample_id)) #2050

#didn't add the sample with duplicate sample_id above
nrow(species_reconc[which(species_reconc$LS.sample_id == "44682364"),])
nrow(species_reconc[which(species_reconc$BS.plant_id == "44226204"),])

#check for duplicate sample_ids
#species_reconc[which(species_reconc$LS.sample_id[which(duplicated(species_reconc$LS.sample_id) == TRUE)][1] %in% status_matched_meta$LS.sample_id),]

#duplicated sample id
species_reconc$LS.sample_id[which(duplicated(species_reconc$LS.sample_id) == TRUE)][1] #41295307

#41295307 sample id
#plant id 40175425
species_reconc[which(species_reconc$LS.sample_id == "41295307"),c(1,2,11,12,21,22,23,24,78,83,84,86,87,95,96,99,129,134,135,137,138,142,158,163,164,176)] #different gps for plants and plant created by different people
species_reconc[which(species_reconc$BS.plant_id == "40175425"),c(1,2,11,12,21,22,23,24,78,83,84,86,87,95,96,99,129,134,135,137,138,142,158,163,164,176)] #I think the issue is two different trees entered into plant database with same plant ID and therefore linked to sample ID because single sample in leaf spectra
status_matched_meta[which(status_matched_meta$LS.sample_id == "41295307"),c(4,9,10,12,13,19,20,21,22,23)]

plants[which(plants$PL.plant_id == "40175425"),c(4,9,10,12,13,17,22,39,40,52,54)] #two plants with same plant ID in plants dataframe

#remove two samples with different trees (for now) - can revisit if plant is fixed
species_reconc <- species_reconc[-which(species_reconc$BS.plant_id == "40175425"),]
nrow(species_reconc)

#check a plant id example
species_reconc[which(species_reconc$BS.plant_id == "8553450"),c(1,2,22,23,84,86,87,95,176)]

unique(species_reconc$BS.plant_id[which(species_reconc$PL.scientific_name == "Betula populifolia Marshall")])
nrow(species_reconc)

#save as reference point
saveRDS(species_reconc, "./data/tidy/species_reconcile_plants.rds")

```

Of 2052 samples of status "submitted", there are 2051 with plant IDs (this is resolved after removing duplicate samples). There are two samples in the leaf spectra dataframe with duplicate sample IDs. These were removed here leaving 2050 samples. After merging the dataframes, two different plants (different species names, different locations) with the same plant ID were identified. These were also removed, leaving 2049 samples. 

Next, the data are tidied to get dates and names into separate fields. The **correct** species names are in the plants database, so in the PL.scientific_name column.

```{r tidy data}

#first, get the field which has correct species names

#make dataframes with structure: sample_ID, proj_name, sp_name.lf, sp_name.pl, date, siteID.ls, plant_id, 
#colnames(species_reconc)
small_matched_meta <- species_reconc %>% 
  select(BS.plant_id, LS.sample_id, LS.project, LS.latitude, LS.longitude, LS.site_id, LS.scientific_name, LS.date_measured, LS.gps_altitude, BS.project, BS.latitude, BS.longitude, BS.scientific_name, BS.site_id, BS.plant_remarks, BS.date_sampled, BS.time_sampled, PL.project, PL.latitude, PL.longitude, PL.site, PL.plot, PL.site_id, PL.plot_id, PL.plant_remarks, PL.vascan_taxon, PL.scientific_name, PL.identification_qualifier, PL.altitude)

str(small_matched_meta)

#remove samples without corresponding plant in plants database 
small_matched_meta <- small_matched_meta[-which(is.na(small_matched_meta$PL.scientific_name)),]

#see which differences there are in species ID in the different datasets
mismatched_names <- data.frame(spectral_name = as.character(), plants_name = as.character(), sample_name = as.character())
for (i in 1:nrow(small_matched_meta)){
  spectral_name <- small_matched_meta$LS.scientific_name[i]
  plants_name <- small_matched_meta$PL.scientific_name[i]
  sample_name <- small_matched_meta$LS.sample_id[i]
  next_one <- data.frame(spectral_name, plants_name, sample_name)
  if (spectral_name != plants_name){
    mismatched_names <- rbind(mismatched_names, next_one)
  }
}

nrow(small_matched_meta)
nrow(mismatched_names)

saveRDS(mismatched_names, "./data/tidy/mismatched_names.rds")

```

There are 167 samples with names that need updating (will update all but those have changes). Spectral dataframe will be updated with the PL.scientific_name column. 

```{r splitting names and dates}

# #convert date to ymd from mdy - not necessary for data from metadata as imported here
# small_matched_meta$YMD <- mdy(small_matched_meta$LS.date_measured)

#separate species names into genus, species, authority

#run for metadata species names (from index page)
names_splitting_metadata <- splitting_species_names(small_matched_meta)

# #check for subspecific names through extra columns
# unique(names_splitting_metadata$extra0)#subsp. var. 
# unique(names_splitting_metadata$extra1)#subsp. var. spissum rugosa strigosus
# unique(names_splitting_metadata$extra2)#angustifolia vaginatum altissima 
# unique(names_splitting_metadata$extra3)#
# unique(names_splitting_metadata$extra4)#subsp. 
# unique(names_splitting_metadata$extra5)#australis
# unique(names_splitting_metadata$extra6)#
# unique(names_splitting_metadata$extra7)
# unique(names_splitting_metadata$extra8)
# unique(names_splitting_metadata$extra9)
# unique(names_splitting_metadata$extra10)
# unique(names_splitting_metadata$extra11)

#those with out specific epithets - leave in for now, except that without plant entry
names_splitting_metadata[which(is.na(names_splitting_metadata$extra0)),c(1:3,7,13,27)] #the quercus entries will be just quercus

#remove the sample without a name in the plant entry
names_splitting_metadata <- names_splitting_metadata[-which(names_splitting_metadata$BS.plant_id == "44667317"),]

#the original entry for that plant without a name
#plants[which(plants$PL.plant_id == "44667317"),]

#remove specific epithet "Linnaeus" for Quercu ssamples
names_splitting_metadata$specific_epithet[which(names_splitting_metadata$specific_epithet == "Linnaeus")] <- NA

#get rid of extra columns due to subspecific names  (edit as needed)
metadata_with_names <- names_splitting_metadata[,-c(30:40)]

#get separate columns for year month day
#is date sampled or date measured most accurate? Are there ever differences?
metadata_with_dates <- metadata_with_names %>%
  mutate(year_sampled = lubridate::year(BS.date_sampled), month_sampled = lubridate::month(BS.date_sampled),
         day_sampled = lubridate::day(BS.date_sampled), year_measured = lubridate::year(LS.date_measured), month_measured = lubridate::month(LS.date_measured), day_measured = lubridate::day(LS.date_measured))

#use presence_absence function to convert locations into separate columns of P/A matrix
#using BS.site_id because PL.site_id has one sample with no site_id #plant_i 39920403
locations_matrix <- as.data.frame(presence_absence(metadata_with_dates$BS.site_id))
#colnames(locations_matrix)

#join locations to date/names meta dataframe
sample_locations_dates <- cbind(metadata_with_dates, locations_matrix)


#check for any other naming errors or inconsistencies 
sort(unique(sample_locations_dates$specific_epithet))

#Eg. pennsylvanica vs pensylvanica vs pensylvanicum
sample_locations_dates[which(sample_locations_dates$specific_epithet == "pennsylvanica"),c(1:3,7,13,27)] #Fraxinus
sample_locations_dates[which(sample_locations_dates$specific_epithet == "pensylvanica"),c(1:3,7,13,27)] #Prunus
sample_locations_dates[which(sample_locations_dates$specific_epithet == "pensylvanicum"),c(1:3,7,13,27)] #Acer

#Do I need to replace _ with space?
sample_locations_dates$subspecific <- gsub("_", " ", sample_locations_dates$subspecific)

saveRDS(sample_locations_dates, "./data/tidy/metadata_with_dates_names_locations.rds")

nrow(sample_locations_dates)

```

I have dates and names parsed and resolved naming issues leaving 2046 samples. I then summarize the number of species and samples by project and by site with dates of measurements.

``` {r summarizing data}

final_metadata <- sample_locations_dates
colnames(final_metadata)

#group by project and species to get number of samples per species per project with P/A of locations per species
#may need to adjust across numbers = just the columns of locations
aggregated_project_data <- final_metadata %>% 
  group_by(LS.project, genus, specific_epithet, subspecific) %>% 
  summarise(n_ind = n(), month_range = paste0(min(month_measured), "-", max(month_measured)),  n_locations = length(unique(BS.site_id)), across(c(35:83), max)) #, n_species = n(unique(species))

#get range of dates and locations by species
aggregated_species_data <- final_metadata %>% 
  group_by(genus, specific_epithet, subspecific) %>% 
  summarise(n_ind = n(), month_range = paste0(min(month_measured), "-", max(month_measured)), n_locations = length(unique(BS.site_id)), across(c(36:84), max)) #, n_species = n(unique(species))

#get number of individuals, species and range of collection by site
aggregated_site_data <- final_metadata %>% 
  group_by(BS.site_id) %>% 
  summarise(n_ind = n(), month_range = paste0(min(month_measured), "-", max(month_measured)), n_species = length(unique(PL.scientific_name))) 

#get info from summary
head(aggregated_project_data)
head(aggregated_species_data)
head(aggregated_site_data)
max(aggregated_species_data$n_ind)
min(aggregated_species_data$n_ind)
max(aggregated_species_data$n_locations)
min(aggregated_species_data$n_locations)
max(aggregated_project_data$n_ind)
min(aggregated_project_data$n_ind)
max(aggregated_site_data$n_species)
min(aggregated_site_data$n_species)
max(aggregated_site_data$n_ind)
min(aggregated_site_data$n_ind)
nrow(aggregated_species_data)

saveRDS(aggregated_site_data, "./data/tidy/aggregated_site_data.rds")
saveRDS(aggregated_project_data, "./data/tidy/aggregated_project_data.rds")
saveRDS(aggregated_species_data, "./data/tidy/aggregated_species_data.rds")

```

There are between 1 and 125 samples per species across all projects, or between 1 and 115 samples per species within one project. Each species is found at between 1 and 10 locations (sites). Sites range from having 1 to 25 species, and between 3 and 268 samples. There are a total of 120 species across all sites and projects.

I will plot the distribution of numbers of samples and species across sites.


``` {r further summary}

# #The objects used:
# aggregated_species_data
# aggregated_project_data
# aggregated_site_data

#get stats on various thresholds of species records and sites
colnames(aggregated_species_data)

#get list of species that are found in >2 locations and where they are located
n2_locations <- aggregated_species_data %>% 
  melt(id = c(1:6)) %>% 
  filter(n_locations > 2 & value == 1) %>% 
  select(c(1:7)) 

#make matrix of just locations where more than 2 species present
n2_locations_matrix <- as.data.frame(presence_absence(n2_locations$variable))

#get locations with the most taxa
top_sites_by_sp <- aggregated_site_data %>% 
  arrange(desc(n_species)) %>% 
  mutate(name = fct_reorder(BS.site_id, desc(n_species)))

#get species with most sites
top_species_by_site <- aggregated_species_data %>% 
  arrange(desc(n_locations)) %>% 
  mutate(name = fct_reorder(c(paste(genus, specific_epithet, subspecific)), desc(n_locations)))

top_species_by_site$name <- gsub(" NA", "", top_species_by_site$name)
  
#plot by site
ggplot(top_sites_by_sp)+
  geom_point(aes(x = name, y = n_species), colour = "blue") +
  geom_point(aes(x = name, y = n_ind/10), colour = "red")

#plot by species
ggplot(top_species_by_site)+
  geom_point(aes(x = name, y = n_locations), colour = "blue") 

length(top_species_by_site$name)
nrow(top_species_by_site)
top_species_by_site$name[1:10]
  
#the top 6 sites by number of species
top_6_sites <- head(top_sites_by_sp)
top_24_sites <- top_sites_by_sp %>% 
  filter(n_species > 3)
top_11_sites <- top_sites_by_sp[1:11,]

#get species lists for top 6 sites
sp_top6_sites <- aggregated_species_data %>% 
  melt(id = c(1:6)) %>% 
  filter(variable %in% top_6_sites$BS.site_id) %>% 
  spread(variable, value) %>% 
  mutate(n_top_sites = rowSums(.[7:12])) %>% 
  filter(n_top_sites > 0) %>% 
  arrange(desc(n_top_sites))

#get species lists for top 24 sites
sp_top24_sites <- aggregated_species_data %>% 
  melt(id = c(1:6)) %>% 
  filter(variable %in% top_24_sites$BS.site_id) %>% 
  spread(variable, value) %>% 
  mutate(n_top_sites = rowSums(.[7:30])) %>% 
  filter(n_top_sites > 0) %>% 
  arrange(desc(n_top_sites, n_ind))

#get species lists for top 11 sites
sp_top11_sites <- aggregated_species_data %>% 
  melt(id = c(1:6)) %>% 
  filter(variable %in% top_11_sites$BS.site_id) %>% 
  spread(variable, value) %>% 
  mutate(n_top_sites = rowSums(.[7:17])) %>% 
  filter(n_top_sites > 0) %>% 
  arrange(desc(n_top_sites))

#get number of species for top sites
nrow(sp_top6_sites)
nrow(sp_top11_sites)
nrow(sp_top24_sites)

#get number of species that are singletons for top sites
nrow(sp_top24_sites[which(sp_top24_sites$n_top_sites == 1),])
nrow(sp_top11_sites[which(sp_top11_sites$n_top_sites == 1),])
nrow(sp_top6_sites[which(sp_top6_sites$n_top_sites == 1),])

#get number of species that are singletons or not for all sites
nrow(aggregated_species_data[which(aggregated_species_data$n_locations == 1),])
nrow(aggregated_species_data[which(aggregated_species_data$n_locations > 1),])
nrow(aggregated_species_data)

#get genera for species found in multiple sites overall
unique(aggregated_species_data$genus[which(aggregated_species_data$n_locations > 1)])

#get number of species for each genus found in multiple sites overall
unique(aggregated_species_data[which(aggregated_species_data$n_locations > 1),]) %>% 
  group_by(genus) %>% 
  summarise(n_sp_per_genus = length(unique(specific_epithet))) %>% 
  arrange(desc(n_sp_per_genus))

```

There are 6, 11, and 24 sites which have high number of species. For the top 6 species, this includes MtMeg-1 (25), CGOP_1 (25), GroboisFieldEL (21), SBLUdeM-f (13), IdB_ile_ste_marg (12), and Mont-Royal (12). These sites have 83 species total, with none that overlap among all 6 sites, and each species found in a maximum of 2 sites of the top 6 by species richness. 61/83 species are found at only a single of the top 6 sites. 

For the top 11 sites, there are 103 species total, with species occurring in maximum of 5/11 sites. 67/103 species are found at only a single of the top 11 sites.

For the top 24 sites, there are 113 species total, with species occurring in a maximum of 6/24 sites. 54/113 species are found at only a single of the top 24 sites. 

For the purpose of comparing spectra across sites by species, there are 62 singleton species (occurring in only 1 site overall), and 58 species which occur in more than 1 site overall, for a total of 120 species. The species that occur in multiple sites belong to 37 genera, including conifers, Rosids, Asterids and monocots, although there are more rosid species overall (based on skimming list) with most species in the genera Acer, Quercus, and Populus. 

``` {r other elements}

#get growth form/habit from flora - not actually included in vascan column - plants$PL.vascan_filter_growth_form

growth_form <- data.frame(species = sort(unique(final_metadata$PL.scientific_name)), growth_form = as.character(NA), stringsAsFactors = FALSE) 

#write to edit and enter growth forms from vascan web search (unable to download)
#those from tropics without growth form so far (NA)
#write.csv(growth_form, "./data/tidy/growth_forms.csv", row.names = FALSE)

#read edited growth_form
growth_form_2 <- read.csv("./data/tidy/growth_forms.csv", header = TRUE, stringsAsFactors = FALSE)

#add column for woody/not
growth_form_2$woody <- NA
growth_form_2$woody[which(growth_form_2$growth_form == "Tree" | growth_form_2$growth_form == "Shrub" | growth_form_2$growth_form == "Tree/Shrub")] <- "1"
growth_form_2$woody[which(growth_form_2$growth_form == "Herb")] <- "0"
growth_form_2$woody[which(growth_form_2$growth_form == "Vine")] <- "1"

#spread to get column per growth form type
spreading_habit <-presence_absence(growth_form_2$growth_form)

#add those in two categories
spreading_habit$Tree[which(spreading_habit$`Tree/Shrub` == 1)] <- 1
spreading_habit$Shrub[which(spreading_habit$`Tree/Shrub` == 1)] <- 1
#get rid of original column
spreading_habit <- spreading_habit[,-4]

#merge types of growth form objects
growth_form_3 <- cbind(scientific_name = growth_form_2$species, spreading_habit, Woody = growth_form_2$woody)
colnames(growth_form_3)[6] <- "No_info"

saveRDS(growth_form_3, "./data/tidy/growth_form_tidy.rds")

#summarize information on growth form
summary_of_growth_forms <- growth_form_3[,-1] %>%  #start with the data frame
  map_dfc(table) %>%  # use map_df from the purrr package to "table" each column
  rownames_to_column("response") %>% #convert the rownames to a column named response
  mutate(resp = case_when(response == 1 ~ "No", #change the resulting 1s to No in resp
                          response == 2 ~ "Yes")) %>%  #change the resulting 2s to Yes in resp
  select(resp, everything(), -response) %>%  #reorder the columns with resp at the front, removing response 
  arrange(desc(resp))

write.csv(summary_of_growth_forms, "./data/tidy/summary_growth_form.csv", row.names = FALSE)

#For taxonomy
#retrieve info from vascan
vascan_output <- vascan_search(growth_form$species, format="json", raw = TRUE)

#convert format from json to list
vascan_r <- fromJSON(vascan_output, TRUE, TRUE)

#get family, order, clade from taxize
#to get higher classification eg Order, family and higher clade

taxonomy_table <- data.frame(species = character(121), clade = character(121), class = character(121), order = character(121), family = character(121), stringsAsFactors = FALSE)

#for simplicity, for Magnoliidae, the class is actually the superorder?
#also excluding the tropical plants
for(i in 1:length(vascan_r$results$matches)){
  if (!is.null(vascan_r$results$matches[[i]]) == TRUE & !is.null(vascan_r$results$matches[[i]]$taxonomicAssertions[[1]]$higherClassification) == TRUE){
    higher_class <- vascan_r$results$matches[[i]]$taxonomicAssertions[[1]]$higherClassification
    highest_clade <- str_split(higher_class, ";")[[1]][1]
    if (str_split(higher_class, ";")[[1]][2] == "Magnoliidae"){
      class <- str_split(higher_class, ";")[[1]][3]
      order <- str_split(higher_class, ";")[[1]][4]
      family <- str_split(higher_class, ";")[[1]][5]
    } else if (str_split(higher_class, ";")[[1]][2] == "Pinidae"){
      class <- str_split(higher_class, ";")[[1]][2]
      order <- str_split(higher_class, ";")[[1]][3]
      family <- str_split(higher_class, ";")[[1]][4]
    }
    species <- vascan_r$results$matches[[i]]$taxonomicAssertions[[1]]$acceptedNameUsage
    taxonomy_table$species[i] <- as.character(species)
    taxonomy_table$clade[i] <- highest_clade
    taxonomy_table$class[i] <- class
    taxonomy_table$order[i] <- order
    taxonomy_table$family[i] <- family
  } else {
    taxonomy_table$species[i] <- NA
    taxonomy_table$clade[i] <- NA
    taxonomy_table$class[i] <- NA
    taxonomy_table$order[i] <- NA
    taxonomy_table$family[i] <- NA
  }
  #return(taxonomy_table)
}

#remove empty rows (tropical species)
taxonomy_table_2 <- taxonomy_table %>% 
  filter(complete.cases(.))

#save it for later
saveRDS(taxonomy_table_2, "./data/tidy/taxonomy_table_higher_class.rds")
write.csv(taxonomy_table_2, "./data/tidy/taxonomy_table_higher_class.csv", row.names = FALSE)
        
# #use if need to subdivide query
# vascan_page <- rbind_pages(vascan_r)

#get environmental variables for gps of plants?

summary_of_growth_forms

#summary of taxonomy
n_per_class <- taxonomy_table_2 %>% 
  group_by(class) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

n_per_order <- taxonomy_table_2 %>% 
  group_by(order) %>% 
  summarize(n = n())%>% 
  arrange(desc(n))

n_per_family <- taxonomy_table_2 %>% 
  group_by(family) %>% 
  summarize(n = n())%>% 
  arrange(desc(n))

#how many globally? how many in Canada?
colnames(n_per_order) <- c("group", "CABO")
colnames(n_per_class) <- c("group", "CABO")
CABO <- rbind(n_per_class, n_per_order)

relative_sampling <- data.frame(group = character(25), canada = integer(25), global = integer(25), stringsAsFactors = FALSE)

dir_taxonomy <- list.files("./data/taxonomy", full.names = TRUE)
taxonomy_files <- lapply(dir_taxonomy, read.delim)
for(i in 1:length(taxonomy_files)){
  num_sp <- length(taxonomy_files[which(taxonomy_files[[i]]$Rank == "Species")])
  name <- as.character(taxonomy_files[[i]]$Scientific.name[1])
  relative_sampling$group[i] <- name
  relative_sampling$canada[i] <- num_sp
}

relative_sampling <- merge(relative_sampling, CABO)

relative_sampling <- relative_sampling %>% 
  mutate(CABO_canada_pct = round((CABO/canada*100), 2)) %>% 
  arrange(desc(CABO_canada_pct))

relative_sampling

write.csv(relative_sampling, "./data/tidy/relative_sampling.csv", row.names = FALSE)

```

Using VASCAN, I resolved species names and retrieved higher order classifications for all species in VASCAN (ie not tropical species). One species was also a synonym (Berberis aquifolium Pursh) so I need to go back and fix that before running VASCAN to get higher classification (or at beginning before any summary stuff). 

I still need to get environmental data for plants (if doing that). 

For a summary of growth forms, there are 32 herbs, 26 shrubs, 57 trees, 1 vine, and 15 without growth form info. There are overall 73 woody (tree/shrub) and 32 herbaceous species of those with info. 

There are five classes: Asteranae (19), Caryophyllanae (1), Lilianae (15), Pinidae (11), and Rosanae (57). There are 20 orders (Fagales: 17, Rosales: 17, Poales: 13, Pinales: 10, Sapindales: 10 etc) and 28 families. There is a heavy bias to Rosids overall.

As for how the taxonomic sampling corresponds with Canadian taxonomic diversity, several orders are well represented (eg Sapindales, Fagales, Pinales with 20-30% of species sampled). Others, especially the Asterids, are underrepresented with between 3.5% (Cornales) and 0.75% (Lamiales) species sampled. Monocots are also underrepresented compared to overall taxonomic diversity with 1% of monocots sampled. 

``` {r dates}

#get range of dates and locations by species
aggregated_species_dates <- final_metadata %>% 
  group_by(genus, specific_epithet, subspecific) %>% 
  summarise(n_ind = n(), earliest_month = min(month_measured), latest_month = max(month_measured), month_range = paste0(min(month_measured), "-", max(month_measured)), duration_sampling = (max(month_measured)-min(month_measured)), n_locations = length(unique(BS.site_id)), across(c(36:84), max)) %>% #, n_species = n(unique(species))
  arrange(desc(duration_sampling))

duration3months <- aggregated_species_dates[which(aggregated_species_dates$duration_sampling > 2),c(1:2,7:8)]

duration2months <- aggregated_species_dates[which(aggregated_species_dates$duration_sampling > 1),c(1:2,7:8)]

duration3months
duration2months

#get samples of species with specific durations
time_samples <- final_metadata[which(final_metadata$genus %in% duration2months$genus & final_metadata$specific_epithet %in% duration2months$specific_epithet),]

#plot distribution of samples

time_to_plot <- time_samples[,c(2,28,29,32,31,33:38,8)]

#use date measured (not sampled)
time_to_plot[which(time_to_plot$day_measured != time_to_plot$day_sampled),]

time_to_plot$month_day <- paste0(time_to_plot$month_measured, "-", time_to_plot$day_measured)

time_to_plot$species_name <- paste0(time_to_plot$genus, " ", time_to_plot$specific_epithet)

time_to_plot$day_of_year <- yday(time_to_plot$LS.date_measured)

#plot it
ggplot(time_to_plot)+
  geom_point(aes(x = day_of_year, y = species_name))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#jpeg("./figures/time_sampling.jpg")
ggplot(time_to_plot)+
  geom_point(aes(x = as.factor(day_measured), y = species_name))+
  facet_wrap(vars(month_measured), scales = "free_x")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

saveRDS(time_to_plot, "./data/tidy/time_to_plot.rds")

```

There are 10 species that were sampled over a three month period or longer. There are 25 species that were sampled over a two month period or longer.

Now, I'll match the spectral data to the updated metadata file (including removing the bad samples above and updating the scientific names).

```{r match spectral to metadata}

#combine metadata from spectral data and from metadata files

#make spectral object into dataframe for manipulation
spec_as_df <- as.data.frame(spectral_data, fix_names = "none", metadata = TRUE)

#trim spectra to just those samples in cleaned metadata
spectra_trimmed  <- spec_as_df[which(spec_as_df$sample_id %in% final_metadata$LS.sample_id),]

#add spectra to metadata itself
merged_meta_spectra <- merge(final_metadata, spectra_trimmed, by.x = "LS.sample_id", by.y = "sample_id")
nrow(merged_meta_spectra)

#there are some duplicate spectra for the same sample numbers - check this out
nrow(final_metadata) #2046
nrow(spectra_trimmed) #2050
length(unique(spectra_trimmed$sample_id)) #2046
nrow(spec_as_df) #2255

dup_spectra <- merged_meta_spectra$LS.sample_id[which(duplicated(merged_meta_spectra$LS.sample_id) == TRUE)]

#spectra numbers and sample numbers are duplicated but the spectra values are different - why?
merged_meta_spectra[which(merged_meta_spectra$LS.sample_id %in% dup_spectra),c(1:5,7,8,13,16,19,20,27,88,91:95)]

#add column for specie names in right format
merged_meta_spectra <- merged_meta_spectra %>% 
  unite(species_names, genus, specific_epithet, subspecific, sep = "_", na.rm = TRUE)
  
#make spectra back into spectral object
trimmed_as_spectra <- as_spectra(spectra_trimmed, name_idx = 1, meta_idxs = c(2:4))


#write spectral object as rds (trimmed simple spectra) and dataframe of spectra merged with metadata as csv

saveRDS(trimmed_as_spectra, "./data/tidy/trimmed_spectra_by_meta.rds")
write.csv(merged_meta_spectra, "./data/tidy/merged_meta_spectra_trimmed.csv", row.names = FALSE)


#get list of all species for phylogeny
species_all <- aggregated_species_data %>% 
  unite(species_names, genus, specific_epithet, subspecific, sep = "_", na.rm = TRUE) %>% 
  select(species_names)

write.csv(species_all, "./data/tidy/species_names_for_phylo.csv", row.names = FALSE)


spectral_data
spec_as_df$sample_name[which(spec_as_df$sample_id == 42591114)]

spec_as_df[which(spec_as_df$sample_id == 42591114 |spec_as_df$sample_id == 42594368 |spec_as_df$sample_id == 42594718),]#42594368    42594718 

final_metadata[which(final_metadata$LS.sample_id == 42591114 |final_metadata$LS.sample_id == 42594368 |final_metadata$LS.sample_id == 42594718),]


```

I need to figure out why those eight (four numbers) spectra are duplicated.

